\clearpage
\markright{Acknowledgment}
\markleft{}
\pagestyle{useheadings}
\section*{Acknowledgment}
\addcontentsline{toc}{section}{Acknowledgment}%
\label{sec:acknowledgement}

First and foremost I would like to thank Professor Anna Wienhard to make it possible for me to work on this topic and to advise this thesis. I would also like to thank her for her patience during my work. Next, I would like to express my deep felt thanks to Doctor Valentina Disarlo. She never hesitated to discuss any of my mathematical problems and the remarks she gave me regarding the text of my thesis were priceless. Thank you very much for the many ours you have invested and the great advice you gave me along the way.
I would like to thank Professor Maria Beatrice Pozzetti and Professor Talia Fernós who were kind enough to discuss my open questions concerning the paper~\cite{MR3509968}. Both of them provided me with very good pointers and references to continue my research. Naturally, I would also like to thank my second corrector, Professor Alessandra Iozzi, to agree to reserve the time necessary to work through and evaluate this (not so short) treatise.
Thanks is also due to Doctor Yves de Cornulier, Doctor Mikael de la Salle and Professor Uri Bader for a very enlightening (online) discussion.\footnote{\url{https://mathoverflow.net/questions/286522/conull-subspace-containing-orbit-of-an-ergodically-acting-group}} 
Of course, this thesis would not be in this form without my proofreaders. I would like to thank Fabian Grünig, Sven Grützmacher, Jakob Schnell, Alexander Schubert and Stefan Zentarra for the hard work and the great remarks.
My family, especially my parents, Jochen and Susanne, and my brother Marc, supported me throughout all of my studies, both financially and emotionally. I find it extraordinary that they made it possible for me to spend my time at university as freely as I did. I was backed in whatever decision I took. Thank you so much for giving me this space and trusting me to pursue my own goals.
Last but by no means least, I would like to thank my girlfriend Sabrina Baier. She endured many discussions and explanations concerning this thesis and more often than not she provided me with very useful insight. All the remarks greatly improved the end product you now hold in your hands. Thank you for always lending me an open ear and for your very deep and even more precise feedback.

\clearpage
\pagestyle{useheadings}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../Master"
%%% ispell-local-dictionary: "en_US"
%%% End:
