\section{Group actions on CAT(0) cube complexes and strong group boundaries}
\label{sec:group}

\begin{defin}[(Non-)elementary action]
  A group action \(\Gamma \to \Aut(X)\) is called \emph{elementary}, if there exists a finite orbit of the action on \(X \sqcup \partial_{\sphericalangle}X\). Otherwise the action is called \emph{non-elementary}.
\end{defin}
\todo{important for essential core}

\begin{defin}[Essential action]
  \todo{Depends on the essential core, reference}
  A group action \(\Gamma \to \Aut(X)\) is called \emph{essential}, if the \(\Gamma\)-essential core is the whole space \(X\).
\end{defin}


\begin{defin}[strong \(\Gamma\)-boundary]
  \todo{Look up strong \(Gamma\)-boundary}
\end{defin}


\subsection{Strong \(\Gamma\)-boundaries}
\label{sec:grp-boundary}

\begin{defin}[Amenable group action]
  Let \(E\) be a separable Banach space. Then I denote by \(E^\ast_1\) the unit ball in its dual space. Let \(S\) be a standard Borel space and for each \(s \in S\) consider \(A_s \subset E^\ast_1\) a non-empty convex weak\(\ast\)-compact subspace. Then \((A_s)_{s \in S}\) will be called a \emph{Borel field of compact convex sets} if \(\{(s, \lambda) \mid \lambda \in A_s\}\) is a Borel subset of \(S \times E^\ast_1\).

  Let \(\Gamma\) be a locally compact group with a measure class preserving group action on \(S\). Furthermore, let \(M\) be a Borel group. Then \(\alpha \colon \Gamma \times s \to M\) is called a \emph{(left) cocycle} if \(\alpha(gh, s) = \alpha(g, hs) \alpha(h, s)\) for all \(g, h \in \Gamma\) and almost all \(s \in S\).

  Each element \(T\) of \(\Isom(E)\) gives rise to a homeomorphism \(T^\ast\) of \(E^\ast_1\) via \((T^\ast\Phi)(x) \coloneqq \Phi(Tx)\) for every \(x \in E\). Thus every cocycle \(\alpha \colon S \times \Gamma \to \Isom(E)\) gives rise to a cocycle \(\alpha^\ast \colon S \times \Gamma \to \operatorname{Homeo}(E^\ast)\) via \(\alpha^\ast (g, s) = (\alpha(g, s)^{-1})^\ast\). A Borel field \((A_s)_{s \in S}\) is called \emph{\(\alpha\)-invariant} if \(\alpha^\ast(g, s) A_{s} = A_{gs}\) for each \(g \in \Gamma\) and almost all \(s \in S\).
  
  Let \(\Gamma\) be a locally compact group and \(S\) a standard Borel space with measure class preserving Borelian \(\Gamma\)-action. The \(\Gamma\)-action is said to be \emph{amenable} if for every separable Banach space \(E\), every Borelian (left) cocycle \(\alpha \colon S \times G \to \Isom(E)\) and every \(\alpha\)-invariant Borel field \((A_s)_{s \in S}\), there exists a Borel map \(\phi \colon S \to E^\ast_1\) such that \(\phi(s) \in A_s\) for almost all \(s\) and for each \(g \in \Gamma\)\(\alpha^\ast(g, s) \phi(s) = \phi(gs)\) almost everywhere.
\end{defin}

\begin{defin}[Strong \(\Gamma\)-boundary]
  Let \(\Gamma\) be a discrete, countable group. A standard Borel space \(B\) is called a \emph{strong \(\Gamma\)-boundary}, if
  \begin{enumerate}
  \item there is a measurable group action of \(\Gamma\) on \(B\) and this action is amenable and
  \item 
  \end{enumerate}
\end{defin}

\begin{lemma}[{\cite[Thm.\ V.6.6]{Conway}}]
  If \(X\) is a compact metric space. Then \(C(X)\) is separable.
\end{lemma}

\begin{thm}
  Let \(B\) be a strong \(\Gamma\)-boundary and \(X\) a compact metric space with a continuous \(\Gamma\) action. Then there exists a \(\Gamma\)-equivariant measurable map \(\phi \colon B \to \mathcal{P}(X)\), where \(\mathcal{P}(X)\) is the set of all probability measures on \(X\).
\end{thm}

\begin{proof}
  Consider \(C(X)\) the space of continuous functions from \(X\) to \(\R\). Together with the supremum norm this is a Banach space, which (by the previous lemma) is also separable. Furthermore, there exists a group action of \(\Gamma\) onf \(C(X)\) via \((gf)(x) = f(g^{-1}x)\) for each \(g \in \Gamma\), \(f \in C(X)\) and \(x \in X\). This action is clearly via isometries. Also for \(\mu \in \mathcal{P}(X)\) we define the action to be \((g\mu)(A) = \mu(g^{-1} A)\) for everye \(g \in \Gamma\) and \(A \in \Sigma\). Then the dual pairing established in the Riesz-Markow representation theorem yields \(\langle gf, \mu\ \rangle = \langle f, g^{-1} \mu \rangle\) or in other words \(g^\ast = g^{-1}\).
  Next, consider
  \[
    \alpha: \Gamma \times B \to \Isom(C(X)), (g, b) \mapsto g.
  \]
  This is a left cocycle.
  Since \(X\) is compact, we have \(C(X) = C_0(X)\) and thus, using the Riesz-Markow representation theorem, we yield \(C(X)^\ast \cong M_{s}(X)\). By Corollary~\ref{cor:banach-alaoglu} we know that \(\mathcal{P}(X)\) is weak\(\ast\)-compact and contained in the unit ball of \(M_s(X)\). Furthermore, \(\mathcal{P}(X)\) is convex and non-empty (take any normalized dirac measure), thus we can choose \(A_b = \mathcal{P}(X)\) for all \(b \in B\). This is in fact an \(\alpha\)-invariant Borel field. Since \(B\) is a strong \(\Gamma\)-boundary, the \(\Gamma\) action is amenable and we yield a measurable map \(\phi \colon B \to C(X)^\ast_1\) such that \(\phi(b) \in A_b = \mathcal{P}(X)\), i.\,e.\ \(\phi \colon B \to \mathcal{P}(X)\) (which is still measuralbe). Lastly, we have
  \begin{align*}
    \phi(gb) & = \alpha^\ast(g, b) \phi(b)\\
               & = \left(\alpha(g,b)^{-1}\right)^\ast \phi(b)\\
               & = \left ( g^{-1}\right)^\ast \phi(b)\\
               & = g\phi(b).
  \end{align*}
\end{proof}

\begin{cor}
  Let \(X\) be a finite dimensional CAT(0) cube complex and \(\bar X\) its Rolle compactification. Let \(\Gamma \to \Aut(X)\) be a discrete countable group acting on \(X\). Furthermore, let \(B\) be a strong \(\Gamma\)-boundary. Then there exists \(\Gamma\)-equivariant measurable map \(\phi\colon B \to \mathcal{P}(\bar X)\), where \(\mathcal{P}(\bar X)\) is the set of probability measures on \(\bar X\).
\end{cor}

\begin{proof}
  By Corollary~\ref{cor:comp-met} it was established that \(\bar X\) is a compact metrizable space. Furthermore, the \(\Gamma\)-action on \(X\) extends to a \(\Gamma\)-action on \(\bar X\).\todo{show that the \(\Gamma\)-action extends to \(\bar X\)} Thus all conditions for the previos theorem are satisfied and we get the desired map \(\phi\colon B \to \mathcal{P}(\bar X)\).
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../Master"
%%% End:
