\section{Introduction}
\label{sec:introduction}

Geometric group theory is a relatively new field in mathematics. One starting point might be seen in Felix Klein's \emph{Erlangen program}~\cite{Klein}. There, he pointed out the deep connection between (abstract) groups and their realization as automorphism groups of topological and geometric spaces. In the following years this point of view was refined and graphs and trees (for example the Cayley graph) were identified as important geometrical objects in order to understand groups from a geometrical point of view. The 1980s and the 1990s were especially fruitful with Gromov's work\ \cite{MR919829} and the progress in Thurston's geometrization program\ \cite{MR648524}. In his treatise Gromov introduced (word-)hyperbolic groups. These groups extend the notion of an isometry group of a hyperbolic space by generalizing hyperbolic geometry to the geometry of non-positively curved spaces, captured in the notion of CAT(0) spaces and considering groups operating on these spaces.

In the 1990s, \textcite{MR1347406} brought certain special CAT(0) spaces, the so-called \emph{CAT(0) cube complexes}, to the attention of a wider audience. General CAT(0) spaces are often too complicated to facilitate the understanding of a group. Instead, cube complexes introduce a combinatorial structure which makes these spaces more \enquote{rigid} and hence, easier to handle. At the same time, they are still flexible enough to admit interesting group actions. In particular, every fundamental group of a surface of genus at least one admits an action on a CAT(0) cube complex (see Examples~\ref{bsp:ccc-torus} and\ \ref{bsp:ccc-genus-2}). Another reason why CAT(0) cube complexes were quickly adopted as a natural object of study is that every tree is a CAT(0) cube complex. Hence, this new object generalized the old \enquote{workhorse} of the field. Interestingly, Sageev was not the first to introduce the complexes. This had already been done by Gromov in his 1987 treatise\ \cite{MR919829}, but only as a particular example of a CAT(0) space.

Another old and very important object class in geometric group theory are the Lie groups. Also in the 1990s \textcite{MR1090825} was able to prove an astounding result which was then coined as \emph{Margulis superrigidity}. It states that under certain (rather weak) conditions a linear representation of a lattice in a Lie group can be extended to the whole group. This superrigidity result had a deep impact and it became an objective to find analogous results in slightly different settings.

In 2016, \textcite{MR3509968} proved a superrigidity theorem for groups acting \emph{essential} and \emph{non-elementary} on CAT(0) cube complexes:

\begin{thm}[{\cite[Theorem 1.5]{MR3509968}}]
  Let \(Y\) be an irreducible finite-dimensional CAT(0) cube complex and \(\Gamma < G_1 \times \dots \times G_l \eqqcolon G\) an irreducible lattice in the product of \(l \geq 2\) locally compact groups. Let \(\Gamma \to \Aut(Y)\) be an essential and non-elementary action on \(Y\). Then the action of \(\Gamma\) on \(Y\) extends continuously to an action of \(G\) by factoring via one of the factors.
\end{thm}

The main ingredient in their result was the construction of a so-called \emph{boundary map}. This map connects the group \(\Gamma\) via a \emph{strong \(\Gamma\)-boundary} \(B\) to the \emph{Roller boundary} \(\partial X\) (a subset of the \emph{Roller compactification} \(\bar X\)) of a CAT(0) cube complex \(X\). More precisely, they proved the existence of a measurable map
\[
  \phi\colon B \to \partial X
\]
which is \(\Gamma\)-equivariant almost everywhere. The space \(B\) is in fact a probability space. The set \(\bar X\) is deeply intertwined with the combinatorial structure of \(X\). Each CAT(0) cube complex has an associated set of hyperplanes \(\mathcal{\hat H}\) dividing the complex in two convex components. The set of all these components \(\mathcal{H}\) is called the set of all \emph{halfspaces} of \(X\). Certain subsets of these sets are denoted \emph{ultrafilters} (see Section~\ref{sec:pocset}). The space \(\bar X\) is then simply given as the set of all ultrafilters (the detailed construction can be found in Chapter~\ref{sec:roller}).

The aim of this thesis is to understand and present the construction of this map (Theorem~\ref{thm:4.1} and Corollary~\ref{cor:4.2}). The idea of the proof is as follows: A strong \(\Gamma\)-boundary \(B\) is defined via two key properties. First, the \(\Gamma\)-action on \(B\) must be amenable. Secondly, the \(\Gamma\)-action must be doubly ergodic with coefficients, which is a strengthening of the standard notion of (double) ergodicity. The group \(\Gamma\) also acts on \(X\) and this action can be extended to the Roller compactification \(\bar X\), where \(\Gamma\) acts via homeomorphisms. The amenability provides a measurable map
\begin{align}
  \psi\colon B \to \mathcal{P}(\bar X)\label{eq:psi}
\end{align}
which is \(\Gamma\)-equivariant almost everywhere and where \(\mathcal{P}(\bar X)\) denotes the set of all regular probability measures on \(\bar X\). The hard part of the proof is then that every probability measure in the image of \(\psi\) identifies a point in \(\bar X\), i.\,e.\ in some sense we would like the probability measures in the image to have a point mass. In order to make this statement precise, we need to introduce \emph{weighted halfspaces} (see Section~\ref{sec:weight}). Let \(\mu\) be a probability measure then the associated weighted halfspaces give the following decomposition of the set of halfspaces
\[
\mathcal{H} = H_\mu^- \sqcup H_\mu \sqcup H_\mu^+.
\]
It turns out that if \(H_\mu = \varnothing\) then \(H_\mu^+\) is an ultrafilter. This would give the desired map from \(\mathcal{P}(\bar X)\) to \(\bar X\). The main work then is to show that \(H_\mu\) indeed vanishes for every \(\mu\) in the image of \(\psi\). For this part to work, we need to introduce further restrictions. We need our complex \(X\) to be \enquote{indecomposable} (i.\,e.\ \emph{irreducible}, see Section~\ref{sec:complex}) and finite-dimensional. Furthermore, the group action on the complex needs to be well-behaved. This is encoded in two properties. First, the group needs to act \emph{essential} which means that \(\Gamma\) needs to be well-behaved with regard to the combinatorial structure of \(X\). Secondly, the group needs to act \emph{non-elementary} which means that \(\Gamma\) needs to be well-behaved with regard to the CAT(0) structure of \(X\). The details of both notions can be found in Section~\ref{sec:special}. With all these definitions in place, we will first be able to show that \(H_\mu\) is always finite. Then, in a second step, we will see that finiteness always implies emptiness. This closes the main proof. As a last step, we will see that the image of \(\phi\) indeed is in the Roller \emph{boundary} not only in the Roller \emph{compactification}. In all the steps of the proof ergodicity (with coefficients) will play a crucial role.

\subsubsection*{Chapter outline}

In Chapter~\ref{sec:complexes}, we will introduce CAT(0) cube complexes. We will start with some metric preliminaries before introducing general CAT(0) spaces. Most important in this early part is the definition of the \emph{visual boundary} (see Definition~\ref{defin:visual}). Afterwards, we will introduce cube complexes and combinatorial maps, which are the structure-preserving mappings (i.\,e.\ morphisms) for our objects. We will give a combinatorial property (Gromov's link condition, see Theorem~\ref{thm:link}) to check the CAT(0) property for cube complexes. Afterwards, we will talk about hyperplanes and halfspaces and some of their important properties (convexity, non-empty intersections, countability).

In Chapter~\ref{sec:roller}, we will first introduce \emph{pocsets}. Pocsets are partially ordered sets admitting a fixed point free, order reversing involution. It turns out that the set of halfspaces of CAT(0) cube complex is always a (discrete) pocset. \textcite{Roller1999} proved the reverse, namely that every (discrete) pocset gives rise to a unique CAT(0) cube complex with this pocset as pocset of halfspaces. The main ingredient of this construction is the notion of an \emph{ultrafilter} which we introduce next. The important observation was that there is a one-to-one correspondence between principal ultrafilters and the vertex set of the CAT(0) cube complex. However, Roller went further and noted that the set of all ultrafilters equipped with a natural topology is a compactification of the vertex set of every CAT(0) cube complex. This lead to the definition of one of our main objects of study: the \emph{Roller compactification} of a CAT(0) cube complex. We will introduce some topological and metric results regarding this space. Afterwards, we will revisit ultrafilters and introduce a second (equivalent) viewpoint, which is more natural for the later arguments. Lastly, we turn towards applications and introduce \emph{intervals} of ultrafilters. These are special (sub-)complexes which have strong restrictions when it comes to sets of halfspaces. In particular, we will see that sets of halfspaces can contain at most finitely many \emph{terminal elements} (i.\,e.\ minimal or maximal elements with regard to the partial order up to going over to the opposite under involution). This property is the main reason we are interested in intervals. We will close the chapter by studying sub-pocsets of halfspaces. We will give conditions under which the associated CAT(0) cube complex can be embedded into the CAT(0) cube complex associated to the actual pocset (Lemma~\ref{lem:lifting}). 

After these two chapters, we will shortly leave the realm of CAT(0) cube complexes and (in Chapter~\ref{sec:measure}) turn towards measure theory and functional analysis. We will start with generalities concerning \emph{measurable spaces, measurable maps} and \emph{(probability) measures}. We will recall the connection between the vector space of continuous functions and the vector space of (signed) measures. However, with these generalities in place, we return to our special case and introduce \emph{weighted halfspaces} (see Definition~\ref{defin:weight}). Lastly, we will prove measurability for certain key maps. 

Up to this point, we did not talk about group actions. This will be remedied in Chapter~\ref{sec:group}. First, we will talk about groups acting on CAT(0) cube complexes. We will see how the group action can be extended to the Roller compactification and we will introduce the notions of \emph{essential} and \emph{non-elementary} group actions citing some important results by \textcite{Caprace2010}. The second half of the chapter is concerned with the introduction of \emph{strong \(\Gamma\)-boundaries} (where \(\Gamma\) is a countable, discrete group). One essential ingredient for this boundary as well as for the proof in general is \emph{ergodicity}. Hence, we have a whole section reserved for this topic. The most important results are:
\begin{itemize}
\item If we have a finite group acting ergodically on a space \(B\), then \(B\) is purely atomic (Lemma~\ref{lem:ergodic-atomic});
\item Ergodicity is inherited by finite index subgroups (Lemma~\ref{lem:4.3}).
\end{itemize}
Afterwards, we will strengthen the notion of ergodicity to \emph{ergodicity with coefficients}. Both notions can be defined via requiring certain \(\Gamma\)-equivariant, measurable maps to be constant. Whereas in the case of ergodicity these maps always have \(\R\) as codomain (with the trivial action by \(\Gamma\)), in the case of ergodicity with coefficients we allow the dual of any separable Banach space that admits a unitary \(\Gamma\)-action. This stronger version of ergodicity leads to a condition for the essentiality of the \(\Gamma\)-action (see Corollary~\ref{cor:4.5}), which we will regularly use. Next, we turn towards \emph{amenable group actions}, which guarantee the existence of certain measurable maps which are \(\Gamma\)-equivariant almost everywhere. With this notion in place, we can define strong \(\Gamma\)-boundaries which are special probability spaces on which \(\Gamma\) acts amenably and ergodically with coefficients. We close the chapter with the result that (thanks to the amenability) we find the map \(\psi\) in Equation~\eqref{eq:psi} (Corollary~\ref{cor:p(x)}).

Chapter~\ref{sec:map} contains the statement and the proof of our main result (Theorem~\ref{thm:4.1} and Corollary~\ref{cor:4.2}). First, we will construct the boundary map using the further assumption that \(H_\mu\) vanishes (as described above). Afterwards, we will prove that if the map exists, it takes values in the Roller boundary. We will then see that \(H_\mu\) being finite already implies it to be empty. The remainder of the chapter builds up the necessary tools to exclude \(H_\mu\) being infinite. The most important results are Proposition~\ref{prop:4.10}, Proposition~\ref{prop:4.21} and Proposition~\ref{prop:f-3.32}. Finally, we descend into the main proof plugging all the previous results together.
\\
\\
As a closing remark, we would like to point out that the material in this thesis by its very nature is very close to the exposition in \textcite{MR3509968}.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../Master"
%%% ispell-local-dictionary: "en_US"
%%% End:
